---
title: "Imputation Study"
author: "Seyed Mostafa Hosseinian"
date: "02/04/2021"
output: html_document
---


# import libraries and packages

```{r echo=TRUE}
library(missForest)
library(mice)
library(faraway)
library(ggplot2)
library(ggpubr)
```


# import data sets

# First data set

```{r echo=TRUE}
data("hprice")
hprice <- hprice[c(-7,-8)]  # Two set of indicators are removed
summary(hprice)
dim(hprice)
```


# Second data set

```{r echo=TRUE}
data(iris)
summary(iris)
dim(iris)
```


# Third data set

```{r echo=TRUE}
data("imports85")
summary(imports85)
dim(imports85)
```



#  1. hprice

## Amputation procedure (10% amputation)

```{r echo=TRUE}
set.seed(0)
hprice.mis1 <- prodNA(hprice, noNA = 0.10)
summary(hprice.mis1)
```


## missForest

```{r echo=TRUE}
set.seed(0)
hprice.imp1 <- missForest(hprice.mis1, verbose = TRUE)
hprice.imp1$OOBerror

hprice.imputed1 <- hprice.imp1$ximp
summary(hprice.imputed1)
```


## MICE

```{r echo=TRUE}
set.seed(0)
hprice.imp1.mice <- mice(hprice.mis1)
hprice.imputed1.mice <- complete(hprice.imp1.mice, action = 1L)

# Imputation Error
(imp.err1.hp <- mixError(hprice.imputed1.mice, hprice.mis1, hprice))
```


## Amputation procedure (20% amputation)

```{r echo=TRUE}
set.seed(0)
hprice.mis2 <- prodNA(hprice, noNA = 0.2)
summary(hprice.mis2)
```


## missForest

```{r echo=TRUE}
set.seed(0)
hprice.imp2 <- missForest(hprice.mis2, verbose = TRUE)
hprice.imp2$OOBerror

hprice.imputed2 <- hprice.imp2$ximp
summary(hprice.imputed2)
```


## MICE

```{r echo=TRUE}
set.seed(0)
hprice.imp2.mice <- mice(hprice.mis2)
hprice.imputed2.mice <- complete(hprice.imp2.mice, action = 1L)

# Imputation Error
(imp.err2.hp <- mixError(hprice.imputed2.mice, hprice.mis2, hprice))
```


## Amputation procedure (30% amputation)

```{r echo=TRUE}
set.seed(0)
hprice.mis3 <- prodNA(hprice, noNA = 0.3)
summary(hprice.mis3)
```


## missForest

```{r echo=TRUE}
set.seed(0)
hprice.imp3 <- missForest(hprice.mis3, verbose = TRUE)
hprice.imp3$OOBerror

hprice.imputed3 <- hprice.imp3$ximp
summary(hprice.imputed3)
```



## MICE

```{r echo=TRUE}
set.seed(0)
hprice.imp3.mice <- mice(hprice.mis3)
hprice.imputed3.mice <- complete(hprice.imp3.mice, action = 1L)

# Imputation Error
(imp.err3.hp <- mixError(hprice.imputed3.mice, hprice.mis3, hprice))
```



# 2. iris

## Amputation procedure (10% amputation)

```{r echo=TRUE}
set.seed(0)
iris.mis1 <- prodNA(iris, noNA = 0.1)
summary(iris.mis1)
```


## missForest

```{r echo=TRUE}
set.seed(0)
iris.imp1 <- missForest(iris.mis1, verbose = TRUE)
iris.imp1$OOBerror

iris.imputed1 <- iris.imp1$ximp
summary(iris.imputed1)
```


## MICE

```{r echo=TRUE}
set.seed(0)
iris.imp1.mice <- mice(iris.mis1)
iris.imputed1.mice <- complete(iris.imp1.mice, action = 1L)

# Imputation Error
(imp.err1.ir <- mixError(iris.imputed1.mice, iris.mis1, iris))
```


## Amputation procedure (20% amputation)

```{r echo=TRUE}
set.seed(0)
iris.mis2 <- prodNA(iris, noNA = 0.2)
summary(iris.mis2)
```


## missForest

```{r echo=TRUE}
set.seed(0)
iris.imp2 <- missForest(iris.mis2, verbose = TRUE)
iris.imp2$OOBerror

iris.imputed2 <- iris.imp2$ximp
summary(iris.imputed2)
```


## MICE

```{r echo=TRUE}
set.seed(0)
iris.imp2.mice <- mice(iris.mis2)
iris.imputed2.mice <- complete(iris.imp2.mice, action = 1L)

# Imputation Error
(imp.err2.ir <- mixError(iris.imputed2.mice, iris.mis2, iris))
```


## Amputation procedure (30% amputation)

```{r echo=TRUE}
set.seed(0)
iris.mis3 <- prodNA(iris, noNA = 0.3)
summary(iris.mis3)
```


## missForest

```{r echo=TRUE}
set.seed(0)
iris.imp3 <- missForest(iris.mis3, verbose = TRUE)
iris.imp3$OOBerror

iris.imputed3 <- iris.imp3$ximp
summary(iris.imputed3)
```


## MICE

```{r echo=TRUE}
set.seed(0)
iris.imp3.mice <- mice(iris.mis3)
iris.imputed3.mice <- complete(iris.imp3.mice, action = 1L)

# Imputation Error
(imp.err3.ir <- mixError(iris.imputed3.mice, iris.mis3, iris))
```



# 3. import85

## Amputation procedure (10% amputation)

```{r echo=TRUE}
set.seed(0)
import.mis1 <- prodNA(imports85, noNA = 0.1)
summary(import.mis1)
```


## missForest

```{r echo=TRUE}
set.seed(0)
import.imp1 <- missForest(import.mis1, verbose = TRUE)
import.imp1$OOBerror

import.imputed1 <- import.imp1$ximp
summary(import.imputed1)
```


## MICE

```{r echo=TRUE}
set.seed(0)
import.imp1.mice <- mice(import.mis1, m = 1)
import.imputed1.mice <- complete(import.imp1.mice, action = 1L)

# Imputation Error
(imp.err1.im <- mixError(import.imputed1.mice, import.mis1, imports85))
```


## Amputation procedure (20% amputation)

```{r echo=TRUE}
set.seed(0)
import.mis2 <- prodNA(imports85, noNA = 0.2)
summary(import.mis2)
```


## missForest

```{r echo=TRUE}
set.seed(0)
import.imp2 <- missForest(import.mis2, verbose = TRUE)
import.imp2$OOBerror

import.imputed2 <- import.imp2$ximp
summary(import.imputed2)
```


## MICE

```{r echo=TRUE}
set.seed(0)
import.imp2.mice <- mice(import.mis2, m = 1)
import.imputed2.mice <- complete(import.imp2.mice, action = 1L)

# Imputation Error
(imp.err2.im <- mixError(import.imputed2.mice, import.mis2, imports85))
```


## Amputation procedure (30% amputation)

```{r echo=TRUE}
set.seed(0)
import.mis3 <- prodNA(imports85, noNA = 0.3)
summary(import.mis3)
```


## missForest

```{r echo=TRUE}
set.seed(0)
import.imp3 <- missForest(import.mis3, verbose = TRUE)
import.imp3$OOBerror

import.imputed3 <- import.imp3$ximp
summary(import.imputed3)
```


## MICE

```{r echo=TRUE}
set.seed(0)
import.imp3.mice <- mice(import.mis3, m = 1)
import.imputed3.mice <- complete(import.imp3.mice, action = 1L)

# Imputation Error
(imp.err3.im <- mixError(import.imputed3.mice, import.mis3, imports85))
```





# Evaluation

## 1. missForest

```{r echo=TRUE}
# NRMSE comparison
values1 <- c(hprice.imp1$OOBerror[1], iris.imp1$OOBerror[1],import.imp1$OOBerror[1],hprice.imp2$OOBerror[1],iris.imp2$OOBerror[1],import.imp2$OOBerror[1],hprice.imp3$OOBerror[1],iris.imp3$OOBerror[1],import.imp3$OOBerror[1])
set1 <- rep(c("hprice", "iris", "import85"), 3)
percentage <- c("10%", "10%", "10%", "20%", "20%", "20%", "30%", "30%", "30%")
data1 <- data.frame(values1, set1, percentage)

values2 <- c(hprice.imp1$OOBerror[2], iris.imp1$OOBerror[2],import.imp1$OOBerror[2],hprice.imp2$OOBerror[2],iris.imp2$OOBerror[2],import.imp2$OOBerror[2],hprice.imp3$OOBerror[2],iris.imp3$OOBerror[2],import.imp3$OOBerror[2])
set2 <- rep(c("hprice", "iris", "import85"), 3)
data2 <- data.frame(values2, set2, percentage)


# plot
p1 <- ggplot(data1, aes(x = set1, y = values1, fill = percentage)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label=round(values1, 4)),vjust=-0.5, size=3, position=position_dodge(0.9)) +
  ylim(c(0,0.2)) +
  xlab("") +
  ylab("NRMSE")

p2 <- ggplot(data2, aes(x = set2, y = values2, fill = percentage)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label=round(values2, 4)),vjust=-0.5, size=3, position=position_dodge(0.9)) +
  ylim(c(0,0.2)) +
  xlab("Data sets") +
  ylab("PFC")

ggarrange(p1, p2, labels = c("NRMSE", "PFC"), ncol = 1, nrow = 2)
```


## 2. missForest vs. MICE

```{r echo=TRUE}
# 10%
values1.compar <- c(hprice.imp1$OOBerror[1],iris.imp1$OOBerror[1],import.imp1$OOBerror[1],imp.err1.hp[1], imp.err1.ir[1],imp.err1.im[1])

df.10perc <- data.frame(values1.compar, sets = rep(c("hprice", "iris", "import85"), 2), alg. = c(rep("missForest",3),rep("MICE", 3)))

# 20%
values2.compar <- c(hprice.imp2$OOBerror[1],iris.imp2$OOBerror[1],import.imp2$OOBerror[1],imp.err2.hp[1], imp.err2.ir[1],imp.err2.im[1])

df.20perc <- data.frame(values2.compar, sets = rep(c("hprice", "iris", "import85"), 2), alg. = c(rep("missForest",3),rep("MICE", 3)))

# 30%
values3.compar <- c(hprice.imp3$OOBerror[1],iris.imp3$OOBerror[1],import.imp3$OOBerror[1],imp.err3.hp[1], imp.err3.ir[1],imp.err3.im[1])

df.30perc <- data.frame(values3.compar, sets = rep(c("hprice", "iris", "import85"), 2), alg. = c(rep("missForest",3),rep("MICE", 3)))



# NRMSE plots
p.10perc <- ggplot(df.10perc, aes(x = sets, y = values1.compar, fill = alg.)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label=round(values1.compar, 4)),
            vjust=-0.5, size=3, position=position_dodge(0.9)) +
  ylim(c(0,0.3)) +
  xlab("") +
  ylab("NRMSE")

p.20perc <- ggplot(df.20perc, aes(x = sets, y = values2.compar, fill = alg.)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label=round(values2.compar, 4)),
            vjust=-0.5, size=3, position=position_dodge(0.9)) +
  ylim(c(0,0.3)) +
  xlab("") +
  ylab("NRMSE")

p.30perc <- ggplot(df.30perc, aes(x = sets, y = values3.compar, fill = alg.)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label=round(values3.compar, 4)),
            vjust=-0.5, size=3, position=position_dodge(0.9)) +
  ylim(c(0,0.3)) +
  xlab("") +
  ylab("NRMSE")

ggarrange(p.10perc, p.20perc, p.30perc, labels = c("10%", "20%", "30%"), ncol = 1, nrow = 3)
```



```{r echo=TRUE}
# 10%
values1.compar.p <- c(hprice.imp1$OOBerror[2],iris.imp1$OOBerror[2],import.imp1$OOBerror[2],imp.err1.hp[2], imp.err1.ir[2],imp.err1.im[2])

df.10perc.p <- data.frame(values1.compar.p, sets = rep(c("hprice", "iris", "import85"), 2), alg. = c(rep("missForest",3),rep("MICE", 3)))

# 20%
values2.compar.p <- c(hprice.imp2$OOBerror[2],iris.imp2$OOBerror[2],import.imp2$OOBerror[2],imp.err2.hp[2], imp.err2.ir[2],imp.err2.im[2])

df.20perc.p <- data.frame(values2.compar.p, sets = rep(c("hprice", "iris", "import85"), 2), alg. = c(rep("missForest",3),rep("MICE", 3)))

# 30%
values3.compar.p <- c(hprice.imp3$OOBerror[2],iris.imp3$OOBerror[2],import.imp3$OOBerror[2],imp.err3.hp[2], imp.err3.ir[2],imp.err3.im[2])

df.30perc.p <- data.frame(values3.compar.p, sets = rep(c("hprice", "iris", "import85"), 2), alg. = c(rep("missForest",3),rep("MICE", 3)))



# PFC plots
p.10perc.p <- ggplot(df.10perc.p, aes(x = sets, y = values1.compar.p, fill = alg.)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label=round(values1.compar.p, 4)),
            vjust=-0.5, size=3, position=position_dodge(0.9)) +
  ylim(c(0,0.4)) +
  xlab("") +
  ylab("PFC")

p.20perc.p <- ggplot(df.20perc, aes(x = sets, y = values2.compar.p, fill = alg.)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label=round(values2.compar.p, 4)),
            vjust=-0.5, size=3, position=position_dodge(0.9)) +
  ylim(c(0,0.4)) +
  xlab("") +
  ylab("PFC")

p.30perc.p <- ggplot(df.30perc, aes(x = sets, y = values3.compar.p, fill = alg.)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label=round(values3.compar.p, 4)),
            vjust=-0.5, size=3, position=position_dodge(0.9)) +
  ylim(c(0,0.4)) +
  xlab("") +
  ylab("PFC")

ggarrange(p.10perc.p, p.20perc.p, p.30perc.p, labels = c("10%", "20%", "30%"), ncol = 1, nrow = 3)
```







